Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-08 edubots_surveys.Rnw
Accept suggestion
Co-authored-by: Francisco Barranco <fbarranco@ugr.es>
Reviewed first discussion paragraphs

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell 
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
the using chatbots to collect course feedback from students in higher
education improved response quality and boosted enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices or teaching strategies based on gamification
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers;
apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the mind of the
surveyed.  \textcolor{black}{In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app based messaging
systems vs. web-based messaging systems integrated in another learning management system).}

This inclusion of university-provided messaging applications in
the survey might explain why there are so few, just 6.7\%, that answer they do not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
people might not use it, but use some extraofficial or opt-in channels such as the other included in the survey; at the end, between these official and extraofficial channels most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant difference. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

@@

The key findings of this study shed light on educators’ preference to use
messaging applications that their institutions support. Technology adoption
literature often \textcolor{black}{focuses} on users’ perceptions of the technology’s usefulness and
ease of use as important prerequisites for successful adoption and
utilization. Nevertheless, in higher education, institutions’ role in
integrating these tools to their educational systems can improve the uptake of
these applications and shape the social and educational experiences of their
students. \textcolor{black}{To accomplish this fruitful integration}, institutions have to ensure that
these messaging applications not only are GDPR (General Data Protection Regulation) compliant to keep students’ data
secure, but also should provide IT support to all stakeholders who use the
applications.

>>>>

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell 
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
the using chatbots to collect course feedback from students in higher
education improved response quality and boosted enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers.
Apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the minds of the
surveyed. In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases, they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app-based messaging
systems vs. web-based messaging systems integrated in another learning management system).

This inclusion of university-provided messaging applications in
the survey might explain why there are so few teachers, just 6.7\%, that answered they did not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
teachers might not use it, but use some extraofficial or opt-in channels such as the others included in the survey; in the end, between these official and extra-official channels, most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant differences. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

@@

The key findings of this study shed light on educators’ preferences to use
messaging applications that their institutions support. Technology adoption
literature often \textcolor{black}{focuses} on users’ perceptions of the technology’s usefulness and
ease of use as important prerequisites for successful adoption and
utilization. Nevertheless, in higher education, institutions’ role in
integrating these tools to their educational systems can improve the uptake of
these applications and shape the social and educational experiences of their
students. \textcolor{black}{To accomplish this fruitful integration}, institutions have to ensure that
these messaging applications not only are GDPR (General Data Protection Regulation) compliant to keep students’ data
secure, but also should provide IT support to all stakeholders who use the
applications.
